---
title: "plot data"
author: "Tao Shi"
date: "4/6/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

# data load
```{r}
require(SummarizedExperiment)
```

```{r}
# load the data by read.table
protein_Group <- read.table("rawdata/proteinGroups.txt", 
                            header= TRUE, 
                            sep = "\t",
                            quote = "", 
                            stringsAsFactors = FALSE)
                          
# sample_id= NA
condition= rep(c ("WT",
                    "WT_Dia",
                    "C182SC277S_Dia",
                    "C182S_Dia"), each= 4) 
replicate= rep (c(1:4), 4)

experiment_design <- data.frame(condition, replicate)  
sample_id <- paste(experiment_design$condition, 
                  experiment_design$replicate, sep = '.')
                           
experiment_design<- data.frame(sample_id, condition, replicate) 

row_data<- protein_Group[, c(1:12, 73:78, 205:209)] 
# by check the colname(protein_Groups) annotations of proteins and peptides
int_matrix <- protein_Group[, c(134:137, 138:141,122:125, 126:129)]
ibaq_matrix<- protein_Group[, c(155: 158, 159: 162, 143: 146, 147: 150)]
lfq_matrix <- protein_Group[, c(175:178, 179: 182, 163:166, 167: 170)]

assays_list <- list( int = int_matrix, ibaq= ibaq_matrix, lfq= lfq_matrix)

assays_list <- lapply(assays_list, function(x){
  rownames(x) <- NULL # reset the row names to seq_len(nrow(x))
  colnames(x) <- NULL
  x <- as.matrix(x) 
  x[x == 0] <- NA
  return(x)
})

SE <- SummarizedExperiment(assays = assays_list, 
                           rowData = row_data, 
                           colData = experiment_design)

save (SE, file= 'rdata/expObj.RData')
```

# require packages
```{r}
require (ggplot2)  
require (UpSetR)  
require (gplots) 
require (corrplot) # plot to visualize matrix correlation
require (PerformanceAnalytics) #an be used to display a chart of a correlation matrix.
require (factoextra)
require (ggpubr) 
require (ggsignif)
require(ggrepel)
cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7") 

```

#  Remove contaminants and reverse matches
```{r}
nrow(SE)
sum(rowData(SE)$Potential.contaminant == '+')
SE <- SE[which(rowData(SE)$Potential.contaminant != '+'), ]

nrow(SE)

sum(rowData(SE)$Reverse == '+')
SE <- SE[which(rowData(SE)$Reverse != '+'),]

nrow(SE)

```
44 potential.contaminant and 33 reverse wwere removed.

# Total proteins identifed per samples
```{r paged.print=FALSE}

counts <- apply(assays(SE)[['lfq']], 2, function(x) sum(!is.na(x))) 
# choose the lfq matrix

plotDf <- as.data.frame(colData(SE))
plotDf$counts <- counts
# plootDf has to be a dataframe

 plotDf$sample_id <- as.character(plotDf$sample_id)
 plotDf$sample_id<- factor(plotDf$sample_id, levels = c ('WT.1','WT.2','WT.3','WT.4',
                                       'WT_Dia.1',  'WT_Dia.2','WT_Dia.3','WT_Dia.4',
                                                 'C182SC277S_Dia.1','C182SC277S_Dia.2', 'C182SC277S_Dia.3', 'C182SC277S_Dia.4',
                'C182S_Dia.1','C182S_Dia.2',  'C182S_Dia.3',  'C182S_Dia.4'   ))   
 
 plotDf$condition<- factor(plotDf$condition, levels = c('WT', 'WT_Dia', 'C182SC277S_Dia', 'C182S_Dia'))    
 
tiff (file= 'plots/Total proteins.tiff', units="in", width =5, height=3, res= 300)
ggplot() + 
  geom_bar(data = plotDf, 
            aes(x = sample_id, y = counts, fill = condition), stat = 'identity') + 
         labs(x = NULL, y= 'Total proteins') + 
  scale_x_discrete(labels= rep(c('R1','R2', 'R3','R4'), 4)) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 60, size = 10, hjust = 1))+
  scale_fill_manual (values = cbPalette) 
  
dev.off()
# The plot is shown in Fig.S5B
```

# PCA assay
```{r }
cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
matrix<- assays(SE)[['lfq']] 

matrix<- log2(matrix) 

matrix[is.na(matrix)] <- 0

colnames(matrix) <- colData(SE)[, 'sample_id']

t(matrix) # transpose a matrix or data.frame

pca <- prcomp(t(matrix))  
# Performs a principal components analysis on the given data matrix and returns the results as an object of class prcomp.

# pca$dim1 captures the direction where the most variation is
fviz_eig(pca) # "ggblot2" based published ready

fviz_pca_ind (pca, axes = c(1, 2), repel = TRUE, col.ind = "cos2")

plotDf <- as.data.frame(pca$x)
plotDf$condition <- colData(SE)[,'condition']

plotDf$condition<- factor(plotDf$condition, levels = c('WT', 'WT_Dia', 'C182SC277S_Dia', 'C182S_Dia'))

tiff(file = 'plots/pca.tiff', units = 'in', height = 3, width = 5, res = 300)
ggplot(data = plotDf) +
  geom_point(aes(x = PC1, y = PC2, color = condition), size = 3) + 
  scale_color_manual(values = cbPalette) + 
  theme_bw()
dev.off()

# The plots is shown as in Fig.5SD
```

# remove the ones with Peptides<3 and save the filtered one
```{r}
sum(rowData(SE)$Peptides < 3)
hist(rowData(SE)$Peptides,xlim = c(0, 30), breaks = 100)
which(rowData(SE)$Peptides > 2)
SE <-SE[which(rowData(SE)$Peptides > 2), ]

nrow(SE)

save(SE, file = 'rdata/SE.RData')


```

# boxplots showing protein distribution

```{r }
load('rdata/SE.RData')
nrow(SE)
assays(SE)[['lfq_log']] <- log2(assays(SE)[['lfq']])

plotDf <- data.frame ( lfq_log = as.vector (assays(SE)[['lfq_log']]),
                     condition = rep(colData(SE)$condition, each = nrow(SE)), 
                     sample = rep(colData(SE)$sample_id, each = nrow(SE)),
                     Gene.names = rep(rowData(SE)$Gene.names, 4))
# as. vector(): converts a distributed matrix into a non-distributed vector
# as.factor (): converts a variable into a factor, but preserves variable and value label attributes
# as.vector to correspond to condition and sample
plotDf$condition<- factor(plotDf$condition, levels= c('WT', 'WT_Dia', 'C182SC277S_Dia', 'C182S_Dia'))
plotDf$sample<- factor(plotDf$sample, levels= c('WT.1', 'WT.2', 'WT.3', 'WT.4', 
                                'WT_Dia.1','WT_Dia.2', 'WT_Dia.3', 'WT_Dia.4',
                                'C182SC277S_Dia.1','C182SC277S_Dia.2', 'C182SC277S_Dia.3', 'C182SC277S_Dia.4',
                                'C182S_Dia.1', 'C182S_Dia.2', 'C182S_Dia.3', 'C182S_Dia.4'))

tiff(file ='plots/boxplot.tiff', units = 'in', height = 3, width = 5, res =300)
ggplot(data = plotDf) + 
  geom_boxplot(aes(x = sample, y = lfq_log, fill = condition)) +
  scale_x_discrete(labels= rep(c('R1','R2', 'R3','R4'), 4)) +
  scale_fill_manual(values = cbPalette) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 60, size = 11, hjust = 1)) + 
  labs(x = NULL, y='Log2(LFQ)')
dev.off()

# the 'boxplot.tiff' is shown in Fig.S5C
# specify TP53

require(ggrepel)

spe<-  plotDf[grep("TP53", plotDf$Gene.names), ]
ggplot(data = plotDf) + 
  geom_boxplot(aes(x = sample, y = lfq_log, fill = condition)) +
  scale_x_discrete(labels= rep(c('R1','R2', 'R3','R4'), 4))+
  scale_fill_manual(values = cbPalette) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90)) + 
  labs(x = 'Sample') +
  geom_text_repel ( data = spe, aes (x= sample, y= lfq_log, label= Gene.names))
  
# TP53 is the most abundant protein and shows equal intensity all over the samples, 

```
